Bootstrapping Multi-atlas Hippocampal Segmentation with MAGeT

### 2013-03-19 paper.Rnw
Bland-Altman plot, plus stddev test of X-validation with poor results... booo

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sweave/Knitr init
<<setup, include=FALSE, cache=FALSE>>=
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
    fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE)
library(ggplot2)
library(Hmisc)
library(reshape2)
library(ADNIMERGE)
library(plyr)
options(digits=3)
theme_set(theme_bw())
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label", "reg_method", "subject"))
all_data$k_diff = all_data$k.y - all_data$k.x 
method_labels <- list("Majority Vote" = "majvote", 
                      "NMI Vote" = "nmi", 
                      "Cross-correlation Vote" = "xcorr")
levels(all_data$method.y) <- method_labels
all_data       = merge(all_data, diagnoses)
remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
@
%                                                                              %
%                                                                              %

The

@@

\begin{figure}
<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote"), 
       aes(x=templates.y, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  #stat_summary(fun.y=mean,geom='line',position="dodge",
  #             aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_smooth(method=lm, formula=y~log(x)) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean Kappa between MAGeT brain and multiatlas" ) 
  #coord_cartesian(ylim=c(0,0.07), xlim=c(0.5,10.5))
@
  \caption{Increase in mean Kappa between MAGeT brain and multi-atlas}
  \label{ADNI1_cross_validation_MA_difference}
\end{figure}

@@

\begin{figure}
<<ADN1-Xval-k-by-disease, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
a_20_tmpls = subset(all_data, 
                      reg_method  == "ANTS" &
                      method.y    == "Majority Vote" &
                      templates.y == 20 &
                      atlases %in% c(1,9))
template_library = read.csv('data/a2a_ants/template_library_composition.csv')
template_library$subject = NULL
a_20_tmpls = merge(a_20_tmpls, template_library,
                   by.x=c('timestamp', 'templates.y'), 
                   by.y=c('timestamp', 'templates'))
ggplot(a_20_tmpls, aes(y=k.y, x=templates_CN,colour=DX)) +  # why does k.x look the same?
  facet_grid( . ~ atlases) + 
  geom_smooth(method='lm',fullrange=F)

@
  \caption{Mean kappa vs. proportion of disease category in template library by disease category}
  \label{}
\end{figure}

@@

\todo{show cost (in registrations) / benefit tradec off graph:  show number of
registrations per Kappa?  or hours of manual labour per Kappa?)}

@@

# Now create one data.frame with all the yr1 volume data we have from each datasource
combined = merge(subjects, mb_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, maper_pruned, by.x="Image.Data.ID", by.y="Source", all.x=TRUE)  # only Baseline
combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, ucsd_pruned , by=c("RID", "VISCODE"), all.x =TRUE)              # TODO: this a unique key?
combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, dx          , by=c("RID","VISCODE"), all.x=TRUE)

attach(combined)
totals = data.frame(RID = RID, VISCODE = VISCODE, DX = DX.bl,
                  MAGeT = MAGeT_L + MAGeT_R, 
                  MAPER = MAPER_L + MAPER_R, 
                  SNT   = SNT_L + SNT_R, 
                  UCSD  = UCSD_L + UCSD_R,
                  FS    = FS_L + FS_R, 
                  FSL   = FSL_L + FSL_R)
detach(combined)
complete_totals = subset(totals, !is.na(SNT) & !is.na(FSL) & !is.na(FS) & !is.na(MAPER) & !is.na(MAGeT))
@

\begin{figure}[h]
<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T>>=
qplot(DX,value,data=melt(complete_totals,measure.vars=c("FS", "FSL","UCSD", "MAPER","MAGeT","SNT")),
      colour=variable,geom="boxplot") + 
      xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
@
  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-scr-volumes-boxplot}
\end{figure}

\begin{figure}[h]
<<ADNI-scr-histogram,dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
melted = melt(complete_totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
ggplot(data=melted, aes(x=value)) + 
       geom_histogram(binwidth=250) +
       facet_grid( variable ~ DX) 
@
  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-scr-histogram}
\end{figure}

\begin{figure}[h]
<<ADNI-src-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
by_snt_volumes = melt(totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))

with(by_snt_volumes,
ggplot(data=by_snt_volumes, aes(x=value)) + geom_histogram(binwidth=250) +
    facet_grid( variable ~ DX) 
     )
#     geom_text(aes(
#       label=paste("MAPER R=", round(cor(maper$SNT,maper$value, use="complete.obs"), 2), 
#                   ", p<", t.test(maper$SNT,maper$value, paired=T)["p.value"]), 
#       x=1800,y= 9500, hjust=0, size=2)) +
#     geom_text(aes(
#       label=paste("MAGeT  R=", round(cor(maget$SNT,maget$value, use="complete.obs"),2), 
#                   ", p<", t.test(maget$SNT,maget$value, paired=T)["p.value"]), 
#       x=1800,y= 9000, hjust=0, size=2)) +
#   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
#   ylab("Total Hippocampal Volume (Automatic Segmentation)")) 
@
  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-baseline-volumes-plot}
\end{figure}

\begin{figure}[h]
<<ADNI-src-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
by_snt_volumes = melt(totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))

with(by_snt_volumes,
ggplot(data=by_snt_volumes, aes(x=value)) + geom_histogram(binwidth=250) +
    facet_grid( variable ~ DX) 
     )
#     geom_text(aes(
#       label=paste("MAPER R=", round(cor(maper$SNT,maper$value, use="complete.obs"), 2), 
#                   ", p<", t.test(maper$SNT,maper$value, paired=T)["p.value"]), 
#       x=1800,y= 9500, hjust=0, size=2)) +
#     geom_text(aes(
#       label=paste("MAGeT  R=", round(cor(maget$SNT,maget$value, use="complete.obs"),2), 
#                   ", p<", t.test(maget$SNT,maget$value, paired=T)["p.value"]), 
#       x=1800,y= 9000, hjust=0, size=2)) +
#   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
#   ylab("Total Hippocampal Volume (Automatic Segmentation)")) 
@
  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-baseline-volumes-plot}
\end{figure}

<<ADNI-scr-summary-table, echo=F, results="asis",cache=FALSE>>=
melted = melt(totals, measure.vars=c("FSL","MAPER","SNT", "FS", "UCSD"))
cor_by_dx    = by(melted[c("MAGeT", "value")], c(melted['variable'], melted['DX']), 
                  function(x) cor(x$MAGeT,x$value, use="pairwise.complete.obs"))
cor_overall  = by(melted[c("MAGeT", "value")], melted['variable'], 
                  function(x) cor(x$MAGeT,x$value, use="pairwise.complete.obs"))
cor_by_dx = t(rbind(cor_by_dx))
cor_overall = rbind(cor_overall)

vol_by_dx = t(rbind(by(totals$MAGeT, totals$DX, mean)))
colnames(vol_by_dx) <- "Mean Volume"
vol_overall = matrix(mean(totals$MAGeT),dimnames=list("Overall"))

dx      = cbind(vol_by_dx, cor_by_dx)
overall <- cbind(vol_overall, cor_overall)

latex(format.df(rbind(dx, overall), dec=2),
      n.cgroup=c(1,5),cgroup=c("","Volume Correlation"),
      n.rgroup=c(4,1), file="", size="scriptsize")

@

@@

gold_diff     = (gold_0.3mm_total - gold_0.6mm_total)/gold_0.3mm_total
mb_bravo_diff = (gold_0.3mm_total - mb_bravo_total)/gold_0.3mm_total
mb_0.6mm_diff = (gold_0.3mm_total - mb_0.6mm_total)/gold_0.3mm_total

subjects = gold_0.6mm[,1]
mb0.6mm_diff_over_gold_diff = cbind(subjects, mb_0.6mm_diff - mean(gold_diff) , res="diffdff")
gold_diff     = cbind(subjects,gold_diff, res="0.3mm gold & 0.6mm gold")
mb_bravo_diff = cbind(subjects,mb_bravo_diff, res="0.3mm gold & MAGeT-on-BRAVO")
mb_0.6mm_diff = cbind(subjects,mb_0.6mm_diff, res="0.3mm gold & MAGeT-at-0.6mm")

diff = rbind(mb_bravo_diff,mb_0.6mm_diff,gold_diff, mb0.6mm_diff_over_gold_diff)
names(diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")

melted = melt(diff, id.vars=c("subject", "res"))
names(melted) <- c("subject","res", "region","voldiff")
ggplot(melted, aes(y=voldiff,x=region)) +
  geom_boxplot() + facet_grid( . ~ res)  + 
  labs(title="Percent difference in MAGeT volumes from original",
       y="Percent difference in volume", 
       x="Region") 
@

<<WAval-kappa,cache=T>>=
data = read.csv('data/WAval-kappa.csv')
ggplot(data, aes(x=label, y=kappa)) + geom_boxplot(y=kappa)
@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sweave/Knitr init
<<setup, include=FALSE, cache=FALSE>>=
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
    fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE)
library(ggplot2)
library(Hmisc)
library(reshape2)
library(ADNIMERGE)
library(plyr)
options(digits=3)
theme_set(theme_bw())
# bland_altman_plot <- function(dat){
#   dat$diff <- dat$pre - dat$post;
#   
#   return ggplot(dat, aes(pre, diff)) + 
#     geom_hline(yintercept = mean(dat$diff) + c(-2, 0, 2) * sd(dat$diff), 
#                linetype=2, color='brown') + 
#     geom_point() 
# }
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label", "reg_method", "subject"))
all_data$k_diff = all_data$k.y - all_data$k.x 
method_labels <- list("Majority Vote" = "majvote", 
                      "NMI Vote" = "nmi", 
                      "Cross-correlation Vote" = "xcorr")
levels(all_data$method.y) <- method_labels
all_data       = merge(all_data, diagnoses)
remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
@
%                                                                              %
%                                                                              %
 % Bland-Altman plot for MA and MAGeT

@@

\begin{figure}
<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote"), 
       aes(x=templates.y, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  coord_cartesian(ylim=c(-0.1,0.05)) + 
  scale_x_continuous(breaks=1:20) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(legend=FALSE) + 
  scale_linetype_discrete(legend=FALSE) + 
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean Kappa between MAGeT brain and multiatlas" ) 
@
  \caption{Increase in mean Kappa between MAGeT brain and multi-atlas}
  \label{ADNI1_cross_validation_MA_difference}
\end{figure}

\begin{figure}
<<ADN1-Xval-k-sd-t-test-p-plot, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
    ANTSmajvote = subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote")

stddev = ddply(ANTSmajvote, c("subject", "label","atlases","templates.y"), function (df) {
  data.frame(
    MA = sd(df$k.x),
    MB = sd(df$k.y) 
  )
})

stddev.t.test = ddply(stddev, c("atlases","templates.y"), function (df) {
  t = t.test(df$MA, df$MB)
  data.frame(
    statistic = t$statistic, 
    p.value = t$p.value
  )
})

ggplot(stddev.t.test, 
       aes(x=templates.y, y=p.value, colour=as.factor(atlases))) + 
  geom_line() + 
  geom_hline(aes(yintercept=0.05), linetype='dashed') + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  scale_x_continuous(breaks=1:20) + scale_y_continuous(breaks=seq(0,1,by=0.05)) + 
  scale_y_log10() + 
  xlab( "Number of Templates" ) + 
  ylab( "p-value" ) 
@
\end{figure}

%- show variance in MAGeT performance by showing nine plots, each with mean multi-atlas performance plotted.

@@

\begin{figure}
<<ADN1-Xval-ANTS-20tmpls-t-test-sd, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
a_20_tmpls = subset(all_data, 
                      reg_method  == "ANTS" &
                      method.y    == "Majority Vote" &
                      templates.y == 20 &
                      atlases     == 9 & 
                      volume.x < 10000)
stddev = ddply(a_20_tmpls, c("subject", "label"), function (df) {
  data.frame(
    MA = sd(df$k.x),  #MA
    MB = sd(df$k.y)  #MB
  )
})

# TODO: against SNT
melted = melt(stddev, id.vars=c("subject", "label"))
ggplot(melted, aes(x=variable, y=value)) + geom_boxplot()
t.test(stddev$MA, stddev$MB)
@
  \caption{}
  \label{}
\end{figure}

\begin{figure}
<<ADN1-Xval-ANTS-20tmpls-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
a_20_tmpls = subset(all_data, 
                      reg_method  == "ANTS" &
                      method.y    == "Majority Vote" &
                      templates.y == 20 &
                      atlases     == 9 &
                      volume.x < 10000)
avg = ddply(a_20_tmpls, c("subject", "label"), function (df) {
  data.frame(
    pre = mean(df$volume.x),  #MA
    post = mean(df$volume.y)  #MB
  )
})
# TODO: against SNT

#df = data.frame(pre=a_20_tmpls$volume.x, post=a_20_tmpls$volume.y)
bland_altman_plot(avg) + 
  xlab('Mean of MA and MAGeT HC Left/Right Volumes') + 
  ylab('MA - MAGeT Volume') + 
  opts(title = 'Bland-Altman Plot')
@
  \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
  \label{}
\end{figure}

\begin{figure}
<<ADN1-Xval-k-by-disease, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
a_20_tmpls = subset(all_data, 
                      reg_method  == "ANTS" &
                      method.y    == "Majority Vote" &
                      templates.y == 20 &
                      atlases %in% c(1,9))
template_library = read.csv('data/a2a_ants/template_library_composition.csv')
template_library$subject = NULL
a_20_tmpls = merge(a_20_tmpls, template_library,
                   by.x=c('timestamp', 'templates.y'), 
                   by.y=c('timestamp', 'templates'))
ggplot(a_20_tmpls, aes(y=k.y, x=templates_CN,colour=DX)) +  # why does k.x look the same?
  facet_grid( . ~ atlases) + 
  geom_smooth(method='lm',fullrange=F) 
 # add 
@
  \caption{Mean kappa vs. proportion of disease category in template library by disease category}
  \label{}
\end{figure}

@@

\todo{show cost (in registrations) / benefit tradec off graph:  show number of
registrations per Kappa?  or hours of manual labour per Kappa?)}
% discuss run-time for each of MAGet and multi-atlas

@@

# Now create one data.frame with all the yr1 volume data we have from each datasource
combined = merge(subjects, mb_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, maper_pruned, by.x="Image.Data.ID", by.y="Source", all.x=TRUE)  # only Baseline
combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
#combined = merge(combined, ucsd_pruned , by=c("RID", "VISCODE"), all.x =TRUE)              # TODO: this a unique key?
combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, dx          , by=c("RID","VISCODE"), all.x=TRUE)

attach(combined)
totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
                  MAGeT = MAGeT_L + MAGeT_R, 
                  MAPER = MAPER_L + MAPER_R, 
                  SNT   = SNT_L + SNT_R, 
                  #UCSD  = UCSD_L + UCSD_R,
                  FS    = FS_L + FS_R, 
                  FSL   = FSL_L + FSL_R)

detach(combined)
#complete_totals = subset(totals, !is.na(SNT) & !is.na(FSL) & !is.na(FS) & !is.na(MAPER) & !is.na(MAGeT))
totals.complete = na.omit(totals)
@

\begin{figure}[h]
<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T>>=
qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
      colour=variable,geom="boxplot") + 
      xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
@
  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-scr-volumes-boxplot}
\end{figure}

% paired t-test against SNT to show similarity (should be low) 
% Kruskal-Willis against  (should be high?)
% post-hoc power analysis to say we can detect this change
\begin{figure}[h]
<<ADNI-scr-histogram,dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
melted   = melt(totals.complete, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
binwidth = 250

# prepare the normals
grid <- with(melted, seq(min(value), max(value), length = 100))
normaldf <- ddply(melted, c("variable", "DX"), function(df) {
  data.frame( 
    value = grid,
    density = dnorm(grid, mean(df$value), sd(df$value)),
    freq = length(df$value) * dnorm(grid, mean(df$value), sd(df$value)) * binwidth,
    length = length(df$value)
  )
})

testsdf <- ddply(melted, c("variable", "DX"), function(df) {
  shapiro = shapiro.test(df$value)
  data.frame(
    shapiro.w = shapiro$statistic,
    shapiro.p = shapiro$p.value,
    mean = mean(df$value),
    sd = sd(df$value)
  )
})

melted.by.snt = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
snt.tests <- ddply(melted.by.snt, c("variable", "DX"), function(df) {
  ks = ks.test(df$SNT, df$value)
  data.frame( 
    ks.D = ks$statistic,
    ks.p = ks$p.value
  )
})

label_x_pos = 10000
ggplot(data=melted, aes(x=value)) + 
  geom_histogram(binwidth=binwidth, alpha=0.8) + 
  geom_line(aes(y=freq, colour='red'), data=normaldf) +
  geom_text(aes(y=29,x=label_x_pos,hjust=1,size=1, 
                label=paste("list(W ==",signif(shapiro.w, digits=3), ", p == ", 
                            signif(shapiro.p,digits=3), ")")), 
            parse=TRUE, data=testsdf) +
  geom_text(aes(y=25,x=label_x_pos,hjust=1,size=1, 
                label=paste("list(D ==",signif(ks.D, digits=3), ", p == ", 
                            signif(ks.p,digits=3), ")")), 
            parse=TRUE, data=snt.tests) + 
  geom_text(aes(y=21,x=label_x_pos,hjust=1,size=1, 
                label=paste("sigma ==",signif(sd, digits=3))), 
            parse=TRUE, data=testsdf) + 
  geom_text(aes(y=17,x=label_x_pos,hjust=1,size=1, 
                label=paste("mu ==",signif(mean, digits=3))), 
            parse=TRUE,data=testsdf) + 
  xlim(0, 10000) + ylim(0,30) + 
  facet_grid(variable ~ DX) + 
  scale_colour_discrete(legend=FALSE) + 
  scale_size_continuous(legend=FALSE) + 
  labs(title="Total hippocampus volume by disease and measurement method", 
       x=expression(paste("Total hippocampal volume (", mm^2, ")")),
       y="Frequency") + 
  opts(axis.title.x = theme_text(vjust=-0.1))
@
  \caption{{\bf ADNI Baseline cohort.} Comparison of total hippocampal volumes as measured by SNT, FSL, FreeSurfer (FS), MAPER, and MAGeT brain (MAGeT). A fitted normal curve is shown in red. $W$ is the Shapiro-Wilkes test statistic measuring normality of the data (signficance indicates non-normality). $D$ is the Kolmogorov-Smirnov test statistic measuring the goodness-of-fit between the distribution of measured volumes and SNT volumes (significance indicates a difference).}
  \label{ADNI-scr-histogram}
\end{figure}

%

% show as a plot
<<ADNI-scr-summary-table, echo=F, results="asis",cache=FALSE>>=
# a rather dirty way of computing the pairwise correlations and then forming
# a printable table out of them.  Sorry.
melted = melt(totals.complete, measure.vars=c("SNT","FSL", "FS", "MAPER"))
cor_by_dx    = by(melted[c("MAGeT", "value")], c(melted['variable'], melted['DX']), 
                  function(x) cor(x$MAGeT,x$value, use="pairwise.complete.obs"))
cor_overall  = by(melted[c("MAGeT", "value")], melted['variable'], 
                  function(x) cor(x$MAGeT,x$value, use="pairwise.complete.obs"))
cor_by_dx = t(rbind(cor_by_dx))
cor_overall = rbind(cor_overall)

vol_by_dx = t(rbind(by(totals.complete$MAGeT, totals.complete$DX, mean)))
colnames(vol_by_dx) <- "Mean Volume"
vol_overall = matrix(mean(totals.complete$MAGeT),dimnames=list("Overall"))

dx      = cbind(vol_by_dx, cor_by_dx)
overall = cbind(vol_overall, cor_overall)
df      = format.df(rbind(dx, overall), dec=2)

latex(df, title="",
      n.cgroup=c(1,ncol(df)-1),cgroup=c("","Volume Correlation"),
      )#file="", size="scriptsize")
@

@@

# calculate percent error from the gold standard measured at 0.3mm resolution
pe_gold0.6mm = (gold_0.6mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
pe_mbbravo   = (mb_bravo_total - gold_0.3mm_total)/gold_0.3mm_total * 100
pe_mb0.6mm   = (mb_0.6mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100

subjects = gold_0.6mm[,1]
pe_gold0.6mm = cbind(subjects,pe_gold0.6mm, res="Resampling to 0.6mm")
pe_mbbravo   = cbind(subjects,pe_mbbravo,   res="MAGeT-on-BRAVO")
pe_mb0.6mm   = cbind(subjects,pe_mb0.6mm,   res="MAGeT-at-0.6mm")

pe = rbind(pe_mbbravo,pe_mb0.6mm, pe_gold0.6mm)
names(pe) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")

melted = melt(pe, id.vars=c("subject", "res"))
names(melted) <- c("subject","res", "region","percenterr")
ggplot(melted, aes(y=percenterr,x=region)) +
  geom_boxplot(aes(colour=res)) +
  labs(title="Percent error in volumes under different measurement regimes",
       y="Percent error in volume", 
       x="Region") 
@

<<WAval-error,cach=T>>=
# compare % error between MAGeT and resampling
mb0.6mm_diff_over_gold_diff = cbind(subjects, abs(gold_diff) - abs(mb_0.6mm_diff), res="diffdff")
names(mb0.6mm_diff_over_gold_diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
melted = melt(mb0.6mm_diff_over_gold_diff, id.vars=c("subject", "res"))
names(melted) <- c("subject","res", "region","voldiff")
ggplot(melted, aes(y=voldiff,x=region)) +
  geom_boxplot() +
  labs(title="Absolute difference in % difference between MAGeT and resampling",
       y="Absolute difference in % difference in volume", 
       x="Region")
@
<<WAval-kappa,cache=T>>=
data = read.csv('data/WAval-kappa.csv')
ggplot(data, aes(x=label, y=kappa)) + geom_boxplot(y=kappa)
@
